Mergeable Types

### 2017-02-28 paper/meta.tex
meta

We now formalize the aforementioned intuitions via a series of
definitions, and theorems.

\begin{definition} [\bfseries Internal and External Ancestors]
Given a branch $b$ and a version $v\in b$, an internal ancestor of $v$
is an ancestor from the same branch $b$. An external ancestor of $v$
is an ancestor from a different branch $b'\neq b$. 
\end{definition}

\begin{definition} [\bfseries External Locus]
Given a branch $b$ and a version $v\in b$, external locus ($v_o$) of
$v$ is an external ancestor that is not an ancestor of any other
external ancestor of $v$. That is, $\under{H}{v_o \preceq v}$, and
there does not exist a $v_o' \not\in b$ such that $\under{H}{v_o'
\preceq v}$ and $\under{H}{v_o \preceq v_o'}$. 
\end{definition}

Note that internal ancestor relation ($\preceq_i$) is, by definition,
a total order. We later prove that the external ancestor relation
($\preceq_o$) is also a total order. Hence, for any given version $v$,
least internal and external ancestors, if exist, must be unique.  We
call the unique least internal ancestor as its \emph{internal locus},
and the unique least external ancestor version as its \emph{external
locus}.

\begin{definition} [\bfseries Mergeability]
\label{def:mergeability}
Given a history $H$, a version $v_1$ and a version $v_2$ that is not
an ancestor of $v_1$ under $H$, $v_2$ is mergeable into $v_1$ (denoted
$\under{H}{v_2 \mbleto v_1}$) iff $v_1$'s external locus is an
ancestor of $v_2$.
\end{definition}

\begin{lemma} [\bfseries Unique External Locus]
Every version $v$ (except the \C{INIT} version) in the history $H$
generated by the rules in Fig.~\ref{fig:opsem} has a unique external
locus.
\end{lemma}

\begin{lemma} [\bfseries Unique LCA]
Every pair of branches in the history $H$ generated by the rules in
Fig.~\ref{fig:opsem} have a unique least common ancestor. 
\end{lemma}

@@

>>>>

We now formalize the notion of mergeability, and the guarantees
provided by the operational semantics.

\begin{definition} [\bfseries Internal and External Ancestors]
Given a branch $b$ and a version $v\in b$, an internal ancestor
($\preceq_i$) of $v$ is an ancestor from the same branch $b$. An
external ancestor ($\preceq_o$) of $v$ is an ancestor from a different
branch $b'\neq b$. 
\end{definition}

\begin{definition} [\bfseries Locus]
Given a branch $b$ and a version $v\in b$, locus ($v_o$) of $v$ is an
ancestor that is not an ancestor of any other external ancestor of
$v$. That is, $\under{H}{v_o \preceq v}$, and there does not exist a
$v_o' \not\in b$ such that $\under{H}{v_o' \preceq v}$ and
$\under{H}{v_o \preceq v_o'}$. We write $L(v)$ to denote the locus
of version $v$.
\end{definition}

In a legal branching history, every version has a unique
locus\footnote{lemma and proof in the appendix}. Thus, locus of a
version is the maxima of its causal history. As usual, the locus of
the latest version on a branch is called the locus of the branch.

\begin{definition} [\bfseries Mergeability]
\label{def:mergeability}
Given a history $H$, a version $v_1$ and a version $v_2$ that is not
an ancestor of $v_1$ under $H$, $v_2$ is mergeable into $v_1$ (denoted
$\under{H}{v_2 \mbleto v_1}$) if and only if $L(v_1)$ is an ancestor
of $v_2$, and no internal ancestor of $v_1$ that is not also an
ancestor of $v_o$ is an ancestor of $v_2$. 
If $v_1$ and $v_2$ are the latest versions on their respective
branches $b_1$ and $b_2$, we say $b_2$ is mergeable into $b_1$, or
more generally, $b_1$ and $b_2$ are mergeable.  
\end{definition}

The following theorem asserts that every pair of branches deemed
mergeable by the above definition are infact mergeable, i.e., they are
guaranteed to have a unique LCA.

\begin{lemma} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{fig:mergeability}, have a unique least common ancestor.
\end{lemma}

@@

Thus, a pair of branches that are not currently mergeable (as per
Def.~\ref{fig:mergeability}), will become mergeable as the branching
history is extended.
